---
title: "HW1_220B"
author: "Jordan"
date: "1/21/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
library(boot)
```

# 5

## a)
```{r}
n <- 1000
y <- seq(0,10,length=n)
fy <- dexp(y,rate=1)
plot(y,fy,type = "l",lwd=2, col=c("blue"), ylab='f(y)',
     xlab = 'y', main=expression(paste('Exponential Distribution ',lambda==1)))
```
The area underneath the curve is equal to 1 since it represents the probabilities of a random variable falling within a range of values, which when integrated over sums to 1. 

## b)

$P(Y > y) = e^{-\lambda x}$


## c)
```{r}
#need to have the likelihood function of exponential distribution, then generate random vector with mean = 1
exp_likeh <- function(lambda,y){
  ans <- (lambda)*exp(-lambda*y)
  return(ans)
}

y <- 1
lambdas <- seq(0,5,length.out = 100)
lamb.lh <- exp_likeh(lambdas,y)

plot(lambdas,lamb.lh, type='l',col='green',lwd=3,
     ylab='Likelihood',xlab=expression(lambda),
     main=expression(paste('Likelihood of ',lambda,' when y=1')))
abline(v=1,col='red',lty=2)
text(1.2,0,labels=expression(lambda['MLE']),col='red')
```

## d)

Theoretical mean:
\begin{eqnarray}
E[Y] = \int^{\infty}_{0} yf(y)dy \\
E[Y] = \int^{\infty}_{0} y\lambda e^{-\lambda y}dy \\
Let \ v' = \lambda e^{-\lambda y}, \ u = y, 
\ v = -e^{-\lambda y}, \ u' = 1, \ du = dy, \ then: \\
= uv - \int vdu \\ 
= [-ye^{-\lambda y}]^{\infty}_0 - \int-e^{-\lambda y}dy \\
Let \ u = -\lambda y, \ du = -\lambda dy, \ then: \\
= (0 - 0) - \int e^u \frac{du}{\lambda} \\ 
= -[-\frac{e^{-\lambda y}}{\lambda}]^{\infty}_0 \\
E[Y] = \frac{1}{\lambda}
\end{eqnarray}



Theoretical variance:
\begin{eqnarray}
Var[Y] = E[Y^2]-(E[Y])^2 \\
E [Y^2] = \int^{\infty}_0 y^2 \lambda e^{-\lambda x} dy \\ 
Let \ u = y^2 , v' = \lambda e^{-\lambda y}, then: \\ 
= -y^2e^{-\lambda x} - \int -e^{\lambda y}2ydy \\
Let \ u = 2y, v' = -e^{-\lambda x}, then: \\
= -x^2e^{-\lambda y} - \frac{2ye^{-\lambda y}}{\lambda} + \int 2\frac{e^{-\lambda y}}{{\lambda}} \\
= -x^2e^{-\lambda y} - \frac{2ye^{-\lambda y}}{\lambda} - \frac{2e^{-\lambda y}}{{\lambda^2}} \\
= [-e^{-\lambda y}(y^2 + \frac{2y}{\lambda} + \frac{2}{\lambda^2})]^{\infty}_0 \\
\quad \\
E[Y^2] = \frac{2}{\lambda^2} \\
Var[Y] = \frac{2}{\lambda^2} - (\frac{1}{\lambda})^2 \\ 
Var[Y] = \frac{1}{\lambda^2}
\end{eqnarray}


## h)
\begin{eqnarray}
f(y | \lambda) = \lambda exp(-\lambda y) \\ 
= exp[ln(\lambda exp(-\lambda y))] \\
= exp[ln(\lambda) - \lambda y] \\
= exp[-\lambda y + ln(\lambda)] \\
\quad \\
\theta_i = -\lambda \\
b(\theta_i) = -ln(\lambda) = -ln(-\theta_i) \\
a(\phi) = 1 \\
\phi = 1 \\
c(y,\phi) = 0
\end{eqnarray}

# 6)
```{r}
n = 1000
b0 <- 1
b1 <- 0.5
x <- seq(-10,10,length=n)
mu <- inv.logit(b0 + b1*x)
```

## d)

```{r}
plot(mu,logit(mu), xlab = expression(pi[i]), ylab = expression(logit(pi[i])), 
     main = expression(paste(pi[i],' vs ',logit(pi[i]))),
     type = 'l', col='blue', lwd=2)
```

## e)

```{r}
plot(x,mu, xlab = expression(x[i]), ylab = expression(pi[i]), 
     main = expression(paste(x[i],' vs ',pi[i])),
     type = 'l', col='red', lwd=2)
```

## f)

```{r}
set.seed(4)
y <- rbinom(n,1,mu)
summary(glm1 <- glm(y ~ x, family="binomial"))
```
$\beta_0 = 1, \beta_1 = 0.5$

